---
title: "Results"
date: "20-10-2021"
output:
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

This markdown compiles all the results for the paper. 

Some steps cannot be reproduced because data cannot be shared. All other steps can be reproduced. Some intermediate results are too large to fit on GitHub. If steps are sourced in the right order, these intermediate results will be stored locally.

You can also skip the intermediate steps and directly reproduce the figures. Figures are saved in ANALYSIS/RESULTS/figures.

```{r setup, include=FALSE}
# Set up
# setwd('/Users/ssmeele/ownCloud/Simeon/MPI AB/PhD thesis/Chapter IV/Coevolution_of_brain_size_and_longevity_in_parrots/ANALYSIS/CODE/markdowns')
# setwd('ANALYSIS/CODE/markdowns')
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ape)
library(phytools)
library(grid)
library(VennDiagram)
library(dagitty)
library(readxl)
library(rethinking)
library(ggridges)

# Clean environment
rm(list = ls())

# Load functions
.functions = sapply(list.files('../functions', pattern = '*.R', full.names = T), source)

# Rerun steps (optional, set to TRUE)
# Some scripts require data that is not publicly available (run local).
# Other scripts take very long to run and do best on an HPC (run cluster).
if(F) source('../00.0 CLEANING - life expectancy.R', chdir = T) # run local
if(F) source('../00.0 CLEANING - body weight.R', chdir = T) # run local
if(F) source('../01.0 PRE-ANALYSIS - body weight.R', chdir = T) # run cluster
if(F) source('../01.0 PRE-ANALYSIS - brain size.R', chdir = T) # run anywhere
if(F) source('../01.0 PRE-ANALYSIS - AFR.R', chdir = T) # run local
if(F) source('../02.0 DATA - retrieve life expectancies.R', chdir = T) # run local
if(F) source('../02.0 DATA - diet.R', chdir = T) # run anywhere
if(F) source('../02.0 DATA - latitude.R', chdir = T) # run anywhere
if(F) source('../02.0 DATA - developmental time.R', chdir = T) # run anywhere
if(F) source('../02.0 DATA - clutch size.R', chdir = T) # run anywhere
if(F) source('../02.0 DATA - insularity.R', chdir = T) # run anywhere
if(F) source('../02.1 DATA - merge all.R', chdir = T) # run anywhere
if(F) source('../03.0 ANALYSIS - final models.R', chdir = T) # run cluster

# Clean env
rm(list = ls())

# Paths
path_script_data_merge = '../02.1 DATA - merge all.R'
path_models = '../../RESULTS/models'
path_pdf_phylo = '../../RESULTS/figures/phylogentic tree.pdf'
path_pdf_phylo_icons = '../../RESULTS/figures/phylogentic tree - with icons.pdf'
path_pdf_slopes = '../../RESULTS/figures/slopes model 2.pdf'
path_pdf_scatter = '../../RESULTS/figures/brain vs life model 2.pdf'

# Load data and models 
source(path_script_data_merge, chdir = T)
for(file in list.files(path_models, '*RData', full.names = T)) load(file)

# Colours
colours = list(blue = '#039BE5', grey = '#CFD8DC', purple = '#9575CD', green = '#81C784', 
               red = '#EF5350', brown = '#8D6E63', yellow = '#FFB74D', pink = '#F06292')
```

## Phylogenetic tree

This step creates the phylogenetic tree with branches coloured by life expectancy. The icons were later added in Inkscape, so this step will reproduce a barebones version of the final figure. 

```{r make phylogentic tree, eval=FALSE, include=FALSE}
# This code is not run, it's only here to show how the figure is made. The final figures was made in inkscape. 
# Pruning tree
tree.tips = tree$tip.label
tree.tips.or = tree.tips
tree.tips = str_replace(tree.tips.or, "_", " ")
drop.tips = ifelse(tree.tips %in% dat$species[which(!is.na(dat$log_mean_life_exp))], NA, tree.tips.or)
drop.tips = na.omit(drop.tips)
tree = drop.tip(tree, drop.tips)
tree.tips.or = tree$tip.label
tree.tips = str_replace(tree.tips.or, "_", " ")

# The value for the colour of the tree
value = dat$log_mean_life_exp[which(!is.na(dat$log_mean_life_exp))] %>% exp
names(value) = tree.tips.or

# Making the flat tree object and changing the colour ramp
ph = data.frame(value = value)
rownames(ph) = names(value)
svl = as.matrix(ph)[,1]
obj = contMap(plot_tree, svl, outline = FALSE, plot = F)
## what is the length of the current color ramp?
n = length(obj$cols)
## change to blue -> red
obj$cols[1:n] = colorRampPalette(c('#1A237E', '#0D47A1', '#2E86C1', '#5DADE2', 
                                   '#F9E79F',
                                   '#EC7063', '#E74C3C', '#CB4335', '#B03A2E', '#CB4335', '#B71C1C'), 
                                 space = 'Lab')(n)

# The genera
genera = tree.tips %>% strsplit(' ') %>% sapply(`[`, 1)
genera[tree.tips == 'Psittacara acuticaudatus'] = 'one' # get rid of misplaced species
genera[tree.tips == 'Psittacara finschi'] = 'two'
genera[tree.tips == 'Psittacara erythrogenys'] = 'three'
genera[tree.tips == 'Psitteuteles iris'] = 'four'
genera[tree.tips == 'Psitteuteles goldiei'] = 'five'
genera[tree.tips == 'Psitteuteles versicolor'] = 'six'
genera[tree.tips == 'Charmosyna papou'] = 'seven'

# Get the name and the y position of each label
label_data = list()
unique_genera = list()
xstart = list()
xend = list()
for(i in unique(genera)){
  nr = round(length(genera[genera == i])/2)
  what.species.middle = tree.tips[genera == i][nr]
  what.species.start = tree.tips[genera == i][1]
  what.species.end = tree.tips[genera == i][round(length(genera[genera == i]))]
  if(length(which(tree.tips == what.species.middle)) != 0) unique_genera = append(unique_genera, i)
  label_data = append(label_data, which(tree.tips == what.species.middle))
  if(length(which(tree.tips == what.species.middle)) != 0) xstart = append(xstart, which(tree.tips == what.species.start))
  if(length(which(tree.tips == what.species.middle)) != 0) xend = append(xend, which(tree.tips == what.species.end))
}
label_data = unlist(label_data)
unique_genera = unlist(unique_genera)
label_data = data.frame(id = label_data, xstart = unlist(xstart), xend = unlist(xend))
# calculate the ANGLE of the labels
number_of_bar = length(value)
angle = -( - 360 * (label_data$id-0.5) /number_of_bar)# I subtract 0.5 because the letter must have the angle of the centre of the bars. 
# Not extreme right(1) or extreme left (0)
# calculate the alignment of labels: right or left
# If I am on the left part of the plot, my labels have currently an angle < -90
label_data$hjust<-ifelse( angle > 90 & angle < 270, 1, 0)
# flip angle BY to make them readable
label_data$angle<-ifelse(angle > 90 & angle < 270, angle+180, angle)
# Remove labels with only one species
unique_genera = unique_genera[label_data$xend - label_data$xstart > 0]
label_data = label_data[label_data$xend - label_data$xstart > 0,]

# Combining all plots
pdf(path_pdf_phylo, 15, 15)

# Treeplot 
plot(obj, type = 'fan', outline = F, mar = c(10, 5, 5, 10), fsize = c(0.01, 1), legend = F)

# Names around
g = ggplot() +       
  geom_segment(data = label_data, aes(x = xstart, xend = xend, y = 13, yend = 13)) +
  ylim(-110, 25) + 
  xlim(0, 217) + 
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.margin = unit(c(-2, 0.5, 0.5, -2), "cm")
  ) +
  coord_polar(start = pi + pi/2 + pi/length(value) - 0.02, direction = -1) +
  geom_text(data = label_data, aes(x = id + 0.5, y = 15, label = unique_genera, hjust = hjust), 
            color = 'black', alpha = 0.6, size = 5, 
            angle = label_data$angle, inherit.aes = FALSE ) 
print(g, newpage = FALSE)

# Tiny histogram as legend colour
ph$order = 'p'
g = ggplot(ph, aes(x = value, y = order)) +
  geom_density_ridges_gradient(aes(fill = ..x..), scale = 3, size = 0.3) +
  scale_fill_gradientn(colours = c('#1A237E', '#0D47A1', '#2E86C1', '#5DADE2', 
                                   '#F9E79F',
                                   '#EC7063', '#E74C3C', '#CB4335', '#B03A2E', '#CB4335', '#B71C1C')) +
  geom_vline(aes(xintercept = 0)) +
  xlim(0, 35) +
  scale_y_discrete(expand = c(0, 0)) +
  coord_cartesian(clip = 'off') + 
  labs(x = 'Life expectancy') +
  theme(legend.position = 'none',
        plot.background = element_blank(),
        panel.background = element_blank(),
        axis.line.x = element_line(),
        axis.ticks.x = element_line(),
        text = element_text(size = 20),
        axis.text = element_text(size = 20),
        axis.title.x = element_text(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = unit(c(34,0.5,0.25,23), 'cm')) 
print(g, newpage = FALSE)

# Save
dev.off()

```

```{r phylogentic tree from PDF, include=TRUE, fig.align="center", fig.cap=paste('Phylogenetic tree of all species for which enough data was available. Branches are coloured acording to life expectancy (see density plot in bottom right). Genera are named if at they contained at least two species. For a version with all species named see Suplemental Materials. (NOTE: I need to redo the silhouttes from only freely available sources. The idea is to have them all the same style as the species at 10 oclock. For this I just need actual pictures and can make them myself.'), fig.height=7, fig.width=7, echo=FALSE}
knitr::include_graphics(path_pdf_phylo_icons)
```


```{r prepare slopes for text, include=FALSE}
post = extract.samples(m_1_final)
sd_rel_brain = sapply(1:nrow(post$true_brain), function(j) 
  sapply(1:ncol(post$true_brain), function(i){
    mu_brain = post$a_brain[j] + post$b_body_brain[j] * post$true_body[j,i]
    (post$true_brain[j,i] - mu_brain)
  }) %>% sd)
slope_1_dist = sapply(1:length(post$b_brain_life), function(i) post$b_brain_life[i] * sd_rel_brain[i])
slope_1_overlap_0 = length(which(slope_1_dist < 0))/length(slope_1_dist)
slope_1 = mean(slope_1_dist) %>% round(2) %>% formatC(format = 'f', flag='0', digits = 2)
post = extract.samples(m_2_final)
sd_rel_brain = sapply(1:nrow(post$true_brain), function(j) 
  sapply(1:ncol(post$true_brain), function(i){
    mu_brain = post$a_brain[j] + post$b_body_brain[j] * post$true_body[j,i]
    (post$true_brain[j,i] - mu_brain)
  }) %>% sd)
slope_2_dist = sapply(1:length(post$b_brain_life), function(i) post$b_brain_life[i] * sd_rel_brain[i])
slope_2_overlap_0 = length(which(slope_2_dist < 0))/length(slope_2_dist)
slope_2 = mean(slope_2_dist) %>% round(2)
slope_dev = mean(post$b_dev_life)
slope_dev_overlap_0 = length(which(post$b_dev_life < 0))/length(post$b_dev_life)
slope_clutch = mean(post$b_clutch_life)
slope_clutch_overlap_0 = length(which(post$b_clutch_life < 0))/length(post$b_clutch_life)
post = extract.samples(m_3_final)
sd_rel_brain = sapply(1:nrow(post$true_brain), function(j) 
  sapply(1:ncol(post$true_brain), function(i){
    mu_brain = post$a_brain[j] + post$b_body_brain[j] * post$true_body[j,i]
    (post$true_brain[j,i] - mu_brain)
  }) %>% sd)
slope_3_dist = sapply(1:length(post$b_brain_life), function(i) post$b_brain_life[i] * sd_rel_brain[i])
slope_3_overlap_0 = length(which(slope_3_dist < 0))/length(slope_3_dist)
slope_3 = mean(slope_3_dist) %>% round(2)
slope_afr = mean(post$b_afr_life)
slope_afr_overlap_0 = length(which(post$b_afr_life < 0))/length(post$b_afr_life)
```
\

\pagebreak

## Slopes

This chunk plots the slopes for the beta parameters in model 2. 

\
```{r model 2 slopes, echo=FALSE, fig.align="center", fig.cap = paste('Slope estimates for predictor variables and phylogenetic signal in model 2. For insularity the difference between islandic and continental species is shown. Grey density plots and lines are the regularising priors. Coloured areas are the postior densities for the parameters. Black lines are 20 samples of the posterior for the phylogenetic covariance.')}
pdf(path_pdf_slopes, 9, 6)
layout(matrix(c(1,1,1,2,2,3,3,
                1,1,1,2,2,3,3,
                1,1,1,2,2,3,3,
                1,1,1,4,4,5,5,
                1,1,1,4,4,5,5,
                6,6,6,4,4,5,5,
                6,6,6,7,7,8,8,
                6,6,6,7,7,8,8,
                6,6,6,7,7,8,8), nrow = 9, ncol = 7, byrow = TRUE))
post = extract.samples(m_2_final)
prior = rnorm(1e6, 0, 0.5) %>% density
mar_reduced_left = c(1, 3, 2, 0)
mar_reduced_right = c(1, 2, 2, 1)
par(oma = c(4, 0, 0, 0))
# Brain
par(mar = c(5.1, 4.1, 2, 2.1))
sd_rel_brain = sapply(1:nrow(post$true_brain), function(j) 
  sapply(1:ncol(post$true_brain), function(i){
    mu_brain = post$a_brain[j] + post$b_body_brain[j] * post$true_body[j,i]
    (post$true_brain[j,i] - mu_brain)
  }) %>% sd)
prior %>% plot(ylim = c(0, 12), xlim = c(-1, 1), main = 'beta relative brain size', xlab = '', ylab = '', cex = 2, cex.axis = 1.25, cex.main = 1.5)
prior %>% polygon(col = alpha(colours[['grey']], 0.5))
sapply(1:length(post$b_brain_life), function(i) post$b_brain_life[i] * sd_rel_brain[i]) %>% density %>% polygon(col = alpha(colours[['purple']], 0.5))
abline(v = 0, lty = 2, lwd = 3, col = alpha('black', 0.5))
mtext('standard deviation', 1, 3, cex = 1)
mtext('density', 2, 3, cex = 1)
text(-0.96, 11.5, 'a)', cex = 1.5, font = 2)
# Body
par(mar = mar_reduced_left)
prior %>% plot(ylim = c(0, 12), xlim = c(-1, 1), main = 'beta body size', xaxt = 'n', yaxt = 'n', xlab = '', ylab = '', cex = 2, cex.main = 1.5)
prior %>% polygon(col = alpha(colours[['grey']], 0.5))
post$b_body_life %>% density %>% polygon(col = alpha(colours[['blue']], 0.5))
abline(v = 0, lty = 2, lwd = 3, col = alpha('black', 0.5))
axis(2)
mtext('density', 2, 3, cex = 0.75)
text(-0.9, 11, 'c)', cex = 1.5, font = 2)
# Insularity
par(mar = mar_reduced_right)
(rnorm(1e6, 0, 0.5) - rnorm(1e6, 0, 0.5)) %>% density %>% 
  plot(ylim = c(0, 12), xlim = c(-1, 1), main = 'diff insularity', xaxt = 'n', yaxt = 'n', xlab = '', ylab = '', cex = 2, cex.main = 1.5)
(rnorm(1e6, 0, 0.5) - rnorm(1e6, 0, 0.5)) %>% density %>% polygon(col = alpha(colours[['grey']], 0.5))
(post$a_life[,1] - post$a_life[,2]) %>% density %>% polygon(col = alpha(colours[['brown']], 0.5))
abline(v = 0, lty = 2, lwd = 3, col = alpha('black', 0.5))
text(-0.9, 11, 'd)', cex = 1.5, font = 2)
# Latitude
par(mar = mar_reduced_left)
prior %>% plot(ylim = c(0, 12), xlim = c(-1, 1), main = 'beta latitude', xaxt = 'n', yaxt = 'n', xlab = '', ylab = '', cex = 2, cex.main = 1.5)
prior %>% polygon(col = alpha(colours[['grey']], 0.5))
post$b_lat_life %>% density %>% polygon(col = alpha(colours[['red']], 0.5))
abline(v = 0, lty = 2, lwd = 3, col = alpha('black', 0.5))
axis(2)
mtext('density', 2, 3, cex = 0.75)
text(-0.9, 11, 'e)', cex = 1.5, font = 2)
# Diet
par(mar = mar_reduced_right)
prior %>% plot(ylim = c(0, 12), xlim = c(-1, 1), main = 'beta diet', xaxt = 'n', yaxt = 'n', xlab = '', ylab = '', cex = 2, cex.main = 1.5)
prior %>% polygon(col = alpha(colours[['grey']], 0.5))
post$b_protein_life %>% density %>% polygon(col = alpha(colours[['green']], 0.5))
abline(v = 0, lty = 2, lwd = 3, col = alpha('black', 0.5))
text(-0.9, 11, 'f)', cex = 1.5, font = 2)
# Phylogenetic signal
par(mar = c(1, 4.1, 2, 2.1))
plot(NULL, xlab = '', ylab = '',
     xlim = c(0,1) , ylim = c(0,1), main = 'phylogenetic signal', cex = 2, cex.axis = 1.25, cex.main = 1.5)
mtext('normalised phylogenetic distance', 1, 3, cex = 1)
mtext('covariance', 2, 3, cex = 1)
x_seq <- seq(from = 0, to = 1, length.out = 1000)
for(i in 1:20) curve( rexp(1, 2)*exp(-rexp(1, 0.1)*x^2) , add = TRUE ,
                      col = alpha(colours[['grey']], 0.4), lwd = 4)
for(i in 1:20){
  curve( post$etasq_life[i]*exp(-post$rhosq_life[i]*x^2) , add = TRUE ,
         col = alpha('black', 0.4), lwd = 4)
}
text(0.01, 0.95, 'b)', cex = 1.5, font = 2)
# Developmental time
par(mar = mar_reduced_left)
prior %>% plot(ylim = c(0, 12), xlim = c(-1, 1), main = 'beta developmental time', xaxt = 'n', yaxt = 'n', xlab = '', ylab = '', cex = 2, cex.main = 1.5)
prior %>% polygon(col = alpha(colours[['grey']], 0.5))
post$b_dev_life %>% density %>% polygon(col = alpha(colours[['yellow']], 0.5))
abline(v = 0, lty = 2, lwd = 3, col = alpha('black', 0.5))
axis(1)
mtext('standard deviation', 1, 3, cex = 0.75)
axis(2)
mtext('density', 2, 3, cex = 0.75)
text(-0.9, 11, 'g)', cex = 1.5, font = 2)
# Clutch size
par(mar = mar_reduced_right)
prior %>% plot(ylim = c(0, 12), xlim = c(-1, 1), main = 'beta clutch size', xaxt = 'n', yaxt = 'n', xlab = '', ylab = '', cex = 2, cex.main = 1.5)
prior %>% polygon(col = alpha(colours[['grey']], 0.5))
post$b_clutch_life %>% density %>% polygon(col = alpha(colours[['yellow']], 0.5))
abline(v = 0, lty = 2, lwd = 3, col = alpha('black', 0.5))
axis(1)
mtext('standard deviation', 1, 3, cex = 0.75)
text(-0.9, 11, 'h)', cex = 1.5, font = 2)
dev.off()

knitr::include_graphics(path_pdf_slopes)
```

\pagebreak

## Scatterplot

This chunk plots the relative brain size vs life expectancy plot to pdf. 

```{r life expectancy and brain size plot model 2, echo = FALSE, fig.align="center", fig.cap = paste('Standardised relative log brain size vs life expectancy for model 2. Pink points represent species where life expectancy was available, pink vertical lines represent the SE for life expectancy. Grey points represent imputed values for life expectancy, grey vertical lines represent the 89% PI for life expectancy. Horizontal lines (pink and grey) represent the 89% PI for standardised relative log brain size. Purple lines represent 20 samples from the posterior for the slope (beta) of the effect of standardised relative log brain size on life expectancy. Standardised relative log brain size is defined in equation X.')}
set.seed(1)
clean.data(c('body', 'brain', 'life', 'lat', 'insularity', 'dev_time', 'diet_protein', 'clutch'))

pdf(path_pdf_scatter, 6, 5)
post = extract.samples(m_2_final)
stand_life = clean_dat$stand_life
se_life = clean_dat$se_life
rel_brain = sapply(1:ncol(post$true_brain), function(i) 
  sapply(1:nrow(post$true_brain), function(j){
    mu_brain = post$a_brain[j] + post$b_body_brain[j] * post$true_body[j,i]
    (post$true_brain[j,i] - mu_brain)
  }))
mean_rel_brain = rel_brain %>% apply(2, mean)
pi_rel_brain = rel_brain %>% apply(2, PI)
plot(NULL, xlim = c(-0.75, 0.75), ylim = c(-3, 3), yaxt = 'n',
     xlab = 'standardised relative log brain size', ylab = 'life expectancy [years]',
     main = '')
axis(2, at = seq(-3, 3, 1), 
     labels = (seq(-3, 3, 1) * sd(dat$log_mean_life_exp, na.rm = T) + 
                 mean(dat$log_mean_life_exp, na.rm = T)) %>% exp %>% round)
for(i in which(!is.na(stand_life)))
  lines(rep(mean_rel_brain[i], 2), stand_life[i] + c(-1, 1) * se_life[i], col = alpha('black', 0.2),
        lwd = 3)
for(i in which(!is.na(stand_life)))
  lines(pi_rel_brain[,i], rep(stand_life[i], 2), col = alpha('black', 0.2),
        lwd = 3)
points(mean_rel_brain, stand_life, pch = 16, col = alpha('black', 0.7), cex = 1.5)
for(i in 1:20) lines(c(-0.75, 0.75), post$b_brain_life[i] * c(-0.75, 0.75),
                     col = alpha(colours[['purple']], 0.5), lwd = 4)
dev.off()

knitr::include_graphics(path_pdf_scatter)
```


